\section{Connections}

\subsection{Practical Rules For How $\nabla$ Acts}

\mybox{
\textbf{Question:} What is the result of the following applications of a torsion free covariant derivative? 
\begin{itemize}
    \item There are some others in the video, but they are basically covered in the lectures so I won't repeat them here. 
    \item $\big(\nabla_{[m} A\big)_{n]}$,
    \item $\big(\nabla_{[m}\omega\big)_{nr]}$.
\end{itemize}
}

\textbf{Solution:} We have 
\bse 
    \begin{split}
        \big(\nabla_{[m} A\big)_{n]} & := \frac{1}{2}\Big[\big(\nabla_mA\big)_n - \big(\nabla_nA\big)_m\Big] \\
        & = \frac{1}{2} \big( A_{n,m} - {\Gamma^r}_{nm}A_r - A_{m,n} + {\Gamma^r}_{mn}A_r\big) \\
        & = A_{[n,m]} + {\Gamma^r}_{[mn]} \\
        & = \frac{1}{2}F_{mn},
    \end{split}
\ese 
where we have used the fact that $\nabla$ is torsion free and so ${\Gamma^a}_{[bc]}=0$.

Next, we have
\bse 
    \big(\nabla_{[m}\omega\big)_{nr]} = \frac{1}{3!}\Big[ \big(\nabla_{m}\omega\big)_{nr} - \big(\nabla_{m}\omega\big)_{rn} + \big(\nabla_{r}\omega\big)_{mn} - \big(\nabla_{r}\omega\big)_{nm} + \big(\nabla_{n}\omega\big)_{rm} - \big(\nabla_{n}\omega\big)_{mr}\Big].
\ese 
If we expand this out we will see that the $\Gamma$s all cancel, e.g.
\bse 
    \begin{split}
        \big(\nabla_{m}\omega\big)_{nr} - \big(\nabla_{n}\omega\big)_{mr} & = \omega_{nr,m} - {\Gamma^s}_{nm}\omega_{sr} - {\Gamma^s}_{rm}\omega_{ns} - \omega_{mr,n} + {\Gamma^s}_{mn}\omega_{sr} + {\Gamma^s}_{rn}\omega_{ms} \\
        & = \omega_{nr,m} - \omega_{mr,n} + 2{\Gamma^s}_{[mn]}\omega_{sr} + 2{\Gamma^s}_{r[n}\omega_{m]s} \\
        & = \omega_{nr,m} - \omega_{mr,n} + 2{\Gamma^s}_{r[n}\omega_{m]s},
    \end{split}
\ese 
and the other $\Gamma$ will cancel with another term's expansion. We are therefore left with 
\bse 
    \big(\nabla_{[m}\omega\big)_{nr]} = \omega_{[nr,m]}.
\ese 

\subsection{Connection Coefficients}

\mybox{
There is a question about stating the chart transformation laws of the $\Gamma$s and stating which class of transformations make the $\Gamma$s look like tensors. We have discussed this in the notes already, but I have but this box here to remind readers to re-read \Cref{rem:GammasTensorTransformation} as it's an important point often missed.
}

\mybox{
\textbf{Question:} Let $(\cM,\cO,\cA,\nabla)$ be the flat plane. Consider two charts that both cover the upper half plane, that is all the points $(a,b)\in\R^2$ with $b>0$, one representing the familiar Cartesian coordinates and the other the familiar polar coordinates on there.  

We already know the chart transition map from Cartesian to polar coordinates is given by 
\bse 
    y\circ x^{-1}(a,b) = \bigg( \sqrt{a^2+b^2}, \arccos\bigg(\frac{a}{\sqrt{a^2+b^2}}\bigg)\bigg),
\ese 
while the inverse transition map from polar to Cartesian is given by 
\bse 
    x\circ y^{-1}(r,\varphi) = \big( r\cos\varphi, r\sin\varphi), \qquad \text{for } r\in\R^+ \text{ and } \varphi\in(0,\pi).
\ese 
Starting from the assumption of vanishing connection coefficient functions in the Cartesian chart, calculate the connection coefficient functions $\Gamma^a_{(y)bc}$ w.r.t. the polar chart!
}

\textbf{Solution}: Recall that the transformation law is 
\bse 
    \Gamma^a_{(y)bc} = \frac{\p y^a}{\p x^k} \frac{\p^2 x^k}{\p y^b\p y^c} + \frac{\p y^a}{\p x^k} \frac{\p x^n}{\p y^b}\frac{\p x^m}{\p y^c} \Gamma^{k}_{(x)mn}.
\ese 
The assumption is $\Gamma^q_{(x)sp}=0$, and so we just need to find the first term. We have 
\bse 
    \bigg(\frac{\p x^k}{\p y^c}\bigg)_{y^{-1}(r,\varphi)} := \p_c\big(x^k\circ y^{-1}\big)(r,\varphi) = \begin{pmatrix}
        \cos\varphi & -r\sin\varphi \\
        \sin\varphi & r\cos\varphi 
    \end{pmatrix}^k_{\,\,\, c}.
\ese 
Now, using the first question in the tangent spaces tutorial above, we have 
\bse 
    \begin{split}
        \bigg(\frac{\p y^a}{\p x^k}\bigg)_p & = \bigg(\frac{\p x^k}{\p y^a}\bigg)^{-1} \\
        & = \frac{1}{r} \begin{pmatrix}
            r\cos\varphi & r\sin\varphi \\
            -\sin\varphi & \cos\varphi 
        \end{pmatrix}^k_{\,\,\, a} \\
        & = \begin{pmatrix}
            \cos\varphi & \sin\varphi \\
            -\frac{1}{r}\sin\varphi & \frac{1}{r}\cos\varphi 
        \end{pmatrix}^k_{\,\,\, a},
    \end{split}
\ese
where we have used the fact that the determinant of the first matrix is $r$.
We can also calculate 
\bse 
    \begin{split}
        \bigg(\frac{\p^2 x^1}{\p y^1 \p y^1}\bigg)_{y^{-1}(r,\varphi)} & = 0 \\
        \bigg(\frac{\p^2 x^1}{\p y^2 \p y^1}\bigg)_{y^{-1}(r,\varphi)} & = -\sin\varphi \\
        \bigg(\frac{\p^2 x^1}{\p y^1 \p y^2}\bigg)_{y^{-1}(r,\varphi)} & = -\sin\varphi \\
        \bigg(\frac{\p^2 x^1}{\p y^2 \p y^2}\bigg)_{y^{-1}(r,\varphi)} & = -r\cos\varphi \\
        \bigg(\frac{\p^2 x^2}{\p y^1 \p y^1}\bigg)_{y^{-1}(r,\varphi)} & = 0 \\
        \bigg(\frac{\p^2 x^2}{\p y^2 \p y^1}\bigg)_{y^{-1}(r,\varphi)} & = \cos\varphi \\
        \bigg(\frac{\p^2 x^2}{\p y^2 \p y^2}\bigg)_{y^{-1}(r,\varphi)} & = -r\sin\varphi.
    \end{split}
\ese 

We then simply have to pick the relevant expressions to find the $\Gamma^a_{(y)bc}$s. For example 
\bse 
    \Gamma^1_{(y)11} = \frac{\p y^1}{\p x^1} \frac{\p^2 x^1}{\p y^1\p y^1} + \frac{\p y^1}{\p x^2} \frac{\p^2 x^2}{\p y^1\p y^1} = 0,
\ese 
and 
\bse
    \begin{split}
        \Gamma^1_{(y)22} & = \frac{\p y^1}{\p x^1} \frac{\p^2 x^1}{\p y^2\p y^2} + \frac{\p y^1}{\p x^2} \frac{\p^2 x^2}{\p y^2\p y^2} \\
        & = \cos\varphi (-r\cos\varphi) + \sin\varphi(-r\sin\varphi) \\
        & = -r(\cos^2\varphi + \sin^2\varphi) \\
        & = -r.
    \end{split}
\ese 

