% Setting section counter to match Penrose Diagrams lecture number.
\setcounter{section}{22}

\section{Diagrams}

\subsection{Penrose Diagram Of A Radiation-Filled Universe}

\mybox{
\textbf{Question}: Find a differential equation for radial null geodesics in a spatially flat FRW universe filled with radiation, using the chart $(t,r,\vartheta,\phi)$ introduced in the lectures. Explicitly write down the precise range of the chart variables. 
}

\textbf{Solution}: A spatially flat FRW universe has $\kappa=0$ and a metric with components 
\bse 
    g_{ab}(t,r,\vartheta,\varphi) = \begin{pmatrix} 
    -1 & 0 & 0 & 0 \\
    0 & a^2(t) & 0 & 0 \\
    0 & 0 & a^2(t)r^2 & 0 \\
    0 & 0 & 0 & a^2(t)r^2\sin^2\vartheta
    \end{pmatrix}_{ab},
\ese
in the chart given. If we have a radiation-filled universe, we have $\omega=-\frac{1}{3}$, and from the last tutorial we have 
\bse 
    a(t) = C \cdot t^{1/2}.
\ese 
Radially null geodesics have $\dot{\vartheta}=0=\dot{\phi}$, so our Lagrangian reads 
\bse 
    0 = -\dot{t}^2 + a^2\cdot \dot{r}^2,
\ese 
which, subbing in our $a(t)$ expression and using the chain rule backwards, gives the differential equation 
\bse 
    \frac{dt}{dr} = \pm C\sqrt{t}.
\ese 

We have seen that in the FRW universe, there is a beginning time (the Big Bang), and so our $t$ coordinate must be lower bounded. We can choose to parameterise it such that $t=0$ is the Big Bang value, giving us the coordinate ranges 
\bse 
    t \in (0,\infty), \qquad r \in (0, \infty), \qquad \vartheta \in (0,\pi), \qand \phi \in (0,2\pi),
\ese
where the $0$ point is removed from $t$'s range as it is not actually a point in our spacetime (it's a singularity). 

\mybox{
\textbf{Question}: Determine the $t$-coordinate of a geodesic in terms of the $r$ coordinate. Draw some of the null geodesics in the underlying chart. 
}

\textbf{Solution}: Solving the differential equation from the previous question gives us 
\bse 
    t_{\pm}(r) = \frac{1}{4}\big(A \pm Cr\big)^2,
\ese
for some integration constant $A$. So we have a series of squared curves, shifted for different values of $A$. 
\begin{center}
    \btik 
        \draw[thick, ->] (0,0) -- (5,0);
        \node at (4.8,-0.2) {$r$};
        \draw[thick, ->] (0,0) -- (0,5);
        \node at (-0.2,4.8) {$t$};
        \begin{scope}
            \clip (0,0) -- (5,0) -- (5,5) -- (0,5) -- (0,0);
            \draw[thick, blue, yshift=-2.25cm] (-3,9) .. controls (0,0) ..  (3,9);
            \draw[thick, blue, yshift=-2.25cm, xshift=1.5cm] (-3,9) .. controls (0,0) ..  (3,9);
            \draw[thick, blue, yshift=-2.25cm, xshift=3cm] (-3,9) .. controls (0,0) ..  (3,9);
            \draw[thick, blue, yshift=-2.25cm, xshift=4.5cm] (-3,9) .. controls (0,0) ..  (3,9);
            \draw[thick, blue, yshift=-2.25cm, xshift=-1.5cm] (-3,9) .. controls (0,0) ..  (3,9);
        \end{scope}
        \draw[fill=white] (0,0) circle [radius=0.06cm];
        \draw[fill=white] (1.5,0) circle [radius=0.06cm];
        \draw[fill=white] (3,0) circle [radius=0.06cm];
        \draw[fill=white] (4.5,0) circle [radius=0.06cm];
        \draw[fill=white] (0,2.4) circle [radius=0.06cm];
    \etik 
\end{center}
The white circles indicated that these points are not part of our diagram as $t$ and $r$ cannot take the value $0$. As these parts are not included, the lines to either side actually represent two separate geodesics, just as we saw with the Schwarzschild drawings in the lectures. 

\mybox{
\textbf{Question}: Find a chart in which the geodesics are lines of constant slope $\pm1$. Determine the range of the coordinates. 
}

\textbf{Solution}: We can rearrange the expression for $t_{\pm}$ in terms of $r$ to give 
\bse 
    r = \pm\frac{2}{C}\sqrt{t_+} - \frac{A}{C}, \qquad r = \mp\frac{2}{C}\sqrt{t_-} + \frac{A}{C}.
\ese 
If we then define 
\bse 
    \Bar{t}_{\pm} := \frac{2}{C}\sqrt{t_{\pm}},
\ese 
we get 
\bse 
    r = \pm \Bar{t}_+ - B, \qquad r = \mp \Bar{t}_- + B,
\ese 
where $B=\frac{A}{C}$. All of these plots are just lines of constant slope $\pm1$. With a bit of thought, its clear that we can just consider either $\bar{t}_+$ or $\bar{t}_-$ and obtain the other results using different values of $B$. So we shall just write 
\bse 
    r_{\pm} = \pm \bar{t} - B.
\ese 
The ranges are 
\bse 
    \bar{t} \in (0,\infty), \qquad r \in (0, \infty), \qquad \vartheta \in (0,\pi), \qand \phi \in (0,2\pi),
\ese 

\mybox{
\textbf{Question}: Choose the so-called null coordinates $u$ and $v$ in which the null geodesics of positive slope are parallel to the $u$-axis and the ones of negative slope are parallel to the $v$-axis. Determine the range of the coordinates. 
}

\textbf{Solution}: We define
\bse 
    u := \bar{t} + r, \qand v := \bar{t}-r,
\ese 
and their ranges are 
\bse 
    u \in (0,\infty), \qand v\in(-\infty,\infty).
\ese 
We then have the conditions $u+v=2\bar{t}>0$ and $u-v=2\bar{r}>0$, so we need to exclude the regions $u<0$, $v+u<0$ and $u-v<0$ from our diagram. We therefore get
\begin{center}
    \btik[scale=0.8]
        \fill[gray!40, opacity=0.8] (0,0) -- (3,3) -- (3,-3) -- (0,0);
        \draw[thick, ->] (-0.2,0) --(3.2,0);
        \node at (3,-0.3) {$u$};
        \draw[thick,->] (0,-3) -- (0,3);
        \node at (-0.3,2.7) {$v$};
        \draw[thick] (0,0) -- (3,3);
        \draw[thick] (0,0) -- (3,-3);
        \begin{scope}
            \clip (0,0) -- (3,3) -- (3,-3) -- (0,0);
            \draw[thick, blue] (0,0.5) -- (3,0.5);
            \draw[thick, blue] (0,1.5) -- (3,1.5);
            \draw[thick, blue] (0,2.5) -- (3,2.5);
            \draw[thick, blue] (0,-0.5) -- (3,-0.5);
            \draw[thick, blue] (0,-1.5) -- (3,-1.5);
            \draw[thick, blue] (0,-2.5) -- (3,-2.5);
            %
            \draw[thick, blue] (0.7,-3) -- (0.7,3);
            \draw[thick, blue] (1.7,-3) -- (1.7,3);
            \draw[thick, blue] (2.7,-3) -- (2.7,3);
        \end{scope}
    \etik 
\end{center}
where the shaded area is the only area we consider. 

The blue lines are the geodesics in this chart. Note that by going to this chart we can compactify without ruining the 90-degree nature of the cone structure. This is exactly why this step is included, and is what \Cref{rem:ConeCompactify} is talking about. 

\mybox{
\textbf{Question}: Compactify, i.e., rescale to finite ranges, each of the two null coordinates by an appropriate transformation. Determine the range of the coordinates. 
}

\textbf{Solution}: We define
\bse 
    p := \arctan(u), \qand q := \arctan(v),
\ese 
which have ranges 
\bse 
    p \in (0, \pi/2), \qand q \in (-\pi/2,\pi/2).
\ese 
Our other range conditions still hold, namely $p+q>0$ and $p-q>0$. 

Our plot looks the same as in the previous question, apart from now the right-hand edge is bounded at value $p=\pi/2$. 

\mybox{
\textbf{Question}: By final transformation, recover the notion of temporal and radial coordinates. Determine the ranges of those coordinates. Draw the Penrose-Carter diagram. 
}

\textbf{Solution}: We define 
\bse 
    T := p+q, \qand R := p-q,
\ese
whose ranges are
\bse 
    T\in (0,\pi), \qand R\in(0,\pi).
\ese 
We have the further constraint $T+R =2U <\pi$, so our Penrose-Carter diagram looks like 
\begin{center}
    \btik 
        \begin{scope}
            \clip[decorate, decoration={snake, segment length=1.5mm, amplitude=0.5mm}] (-0.15,0) -- (5,0) -- (-0.15,5) -- (-0.15,0);
            \clip (0,4.5) -- (0,-1) -- (5.5,-1) -- (0,4.5);
            \fill[gray!40, opacity=0.8] (0,4.5) -- (0,-1) -- (5.5,-1) -- (0,4.5);
            \draw[thick, blue] (0,0) -- (4,4);
            \draw[thick, blue] (0,1.5) -- (3,4.5);
            \draw[thick, blue] (0,3) -- (3,6);
            \draw[thick, blue] (0,-1.5) -- (3,1.5);
            \draw[thick, blue] (0,-3) -- (4,1);
            %
            \draw[thick, blue] (0,3.5) -- (4,-0.5);
            \draw[thick, blue] (0,2) -- (2.5,-0.5);
            \draw[thick, blue] (0,0.5) -- (1,-0.5);
        \end{scope}
        \draw[thick, decorate, decoration={snake, segment length=1.5mm, amplitude=0.5mm}] (0,0) -- (4.5,0);
        \draw[thick] (4.5,0) -- (0,4.5);
        \draw[thick, dashed] (0,0) -- (0,4.5);
    \etik 
\end{center}
We have used the snake-like line to indicate the Big Bang singularity, and a dashed line on the left-hand side to remind us that there is nothing wrong here (i.e. lines that go off to the left just come back form the dashed line --- think about rotating the diagram by reinstating $\phi$).  


