\section{Tutorial}

\subsection{Restricting The Atlas}

\mybox{
Let $(\R,\cO_{\text{standard}})$ be a topological space. Let it be further equipped with an atlas $\cA = \{ (\R,x), (\R,y)\}$ where $x:\R\to\R$; $a\mapsto x(a) = a$ and $y:\R\to\R$; $a\mapsto y(a)= a^3$. 

\textbf{Question}: Construct the chart transition map $y\circ x^{-1}:\R \to \R$ and give its differentiability class. 
}
\textbf{Solution}: We have $x^{-1}:\R \to \R$; $a\mapsto x^{-1}(a) = a$, and so $y\circ x^{-1}:\R\to\R$; $a\mapsto a^3$. This is $C^{\infty}(\R)$. 

\mybox{
\textbf{Question}: Also construct the chart transition map $x\circ y^{-1}:\R\to\R$. Is $(\R,\cO_{\text{standard}},\cA)$ a differentiable manifold?
}

\textbf{Solution}: We have $x\circ y^{-1}: \R\to\R$; $a\mapsto \sqrt[3]{a}$, but this function is not even $C^1(\R\to\R)$ as the first derivative is $\frac{1}{3}a^{-2/3}$, which blows up as $a\to 0$. So no it is not a differentiable manifold. 

\mybox{
\textbf{Question}: Restrict the atlas $\cA$ to an atlas $\widetilde{\cA}$ in order to make $(\R,\cO_{\text{standard}},\widetilde{\cA})$ a smooth manifold.
}
\textbf{Solution}: The problem above came from the chart transition maps. Both chart maps themselves are $C^{\infty}(\R\to\R)$ and their domain is the whole manifold (namley $\R$) and so if we just remove one of the two charts we get a smooth manifold. 

\subsection{Soft Squares on $\R\times\R$}

\mybox{
Let $\cM=\R\times\R$ equipped with the soft square topology $\cO_{ssq}$ and an atlas $\cA=\{(U_n,x_n)\}$, where $U_n = \{(x,y)\in\R\times\R \, | \, |x|<n, |y|<n, n\in \mathbb{N}^+\}$ and 
\bse 
    x_n : U_n \to x_n(U_n) \ss \R^2; \, (x,y) \mapsto x_n\big((x,y)\big) := \bigg(\frac{x+y}{2n}, \frac{x-y}{2n}\bigg).
\ese
\textbf{Question}: Recall the definition of a chart and show that the $(U_n,x_n)$ are indeed charts.
}

\textbf{Solution}: For $(U_n,x_n)$ to be charts, we need to show that $U_n\in \cO_{ssq}$ and that the $x_n$s are homeomorphisms (i.e. inevitable and continuous). 

The soft square topology is rather self explanatory, it is the set of squares around the origin without the boundary. This is exactly the definition of the $U_n$s where the sides of consecutive $U_n$ increase by $2n$. So we have $U_n\in\cO_{ssq}$.

Next we need to show that the chart maps are homeomorphisms. It is clear that $x_n$ is continuous w.r.t. $\cO_{ssq}$ and $\cO_{\text{standard}}$. So we just need to check that $x_n^{-1}$ exists and is continuous. We have 
\bse 
    x_n^{-1}(a,b) = \big(n(a+b), n(a-b)\big),
\ese 
which is again clearly continuous. 

Therefore we know that the $(U_n,x_n)$ are indeed charts. 

\mybox{
\textbf{Question}: Show that $\cA$ is a $C^k$-atlas by explicitly constructing the chart transition maps. What is $k$? 
}
\textbf{Solution}: The chart transition maps are given by 
\bse 
    \begin{split}
        x_m \circ x_n^{-1} : x_n(U_n\cap U_m) & \to x_m(U_n\cap U_m) \\
        (a,b) & \mapsto \bigg(\frac{na}{m}, \frac{nb}{m}\bigg) = \frac{n}{m}(a,b).
    \end{split}
\ese 
This tells us that $x_m\circ x_n^{-1} = \frac{n}{m}\b1_{\R^2}$, which together with $m\neq 0$ tells us $k=\infty$. 

\mybox{
\textbf{Question}: Construct at least one other chart that would lie in the maximal extension of $\cA$ and prove that it does. 
}

\textbf{Solution}: There are many, but we could take the map $\widetilde{x}:U_5 \to \widetilde{U_5}\ss\R^2$; $(a,b) \mapsto (a,b)$, i.e. it is the identity map restricted to $U_5$. It is clear that this will be $C^{\infty}$ compatible with any overlapping charts and so it lies in the atlas. 

\subsection{Undergraduate Multi-Dimensional Analysis}

\mybox{
\textbf{Question}: There is a question about calculating partial derivatives. I am not including it here as its fairly straight forward.
}

\subsection{Differentiability On A Manifold}

\mybox{
\textbf{Question}: There is a question about drawing a diagram to show a bunch of different spaces and maps. It will take a while to draw in Tikz and there are plenty examples in the notes themselves, so I haven't included it here. This is followed by a calculation of a derivative of a map, it is worth seeing this exercise, so please \href{https://www.youtube.com/watch?v=FXPdKxOq1KA&list=PLFeEvEPtX_0RQ1ys-7VIsKlBWz7RX-FaL&index=4}{the video}. 
}

