\section{Tutorial}

\subsection{Where Connection Coefficients Appear}

\mybox{
\textbf{Question}: Determine the coefficients of the Riemann tensor with respect to a chart $(U,x)$ in terms of the connection coefficient functions. 
}

\textbf{Solution}: Recall the definition 
\bse 
    \Riem(\omega,Z,X,Y) := \omega : \Big( \nabla_X\nabla_YZ - \nabla_Y\nabla_XZ - \nabla_{[X,Y]}Z\Big).
\ese 
We have already seen\footnote{Or more correctly, you have shown as an exercise in Lecture 8.} that Riem is $C^{\infty}$-linear in all its entries and so we can just consider basis elements. That is, we can set (using the notation $\p_j := \frac{\p}{\p x^j}$)
\bse 
    \omega = dx^i, \qquad Z = \p_j, \qquad X = \p_k, \qand Y = \p_m.
\ese 
So we have 
\bse 
    \begin{split}
        \Riem(dx^i,\p_j,\p_k,\p_m) & := dx^i : \Big(\nabla_k\nabla_m\p_j - \nabla_m\nabla_k\p_j - \nabla_{[\p_k,\p_m]}\p_j\Big) \\ 
        & = dx^i : \Big[ \nabla_k \big(\Gamma^r_{(x)jm}\p_r\big) - \nabla_m\big( \Gamma^r_{(x)jk}\p_r\big) \Big] \\
        & = dx^i : \Big[ \Gamma^r_{(x)jm,k}\p_r + \Gamma^r_{(x)jm}\Gamma^s_{(x)rk}\p_s - \Gamma^r_{(x)jk,m}\p_r - \Gamma^r_{(x)jk}\Gamma^s_{(x)rm}\p_s \Big] \\
        {\Riem^i}_{jkm} & = \Gamma^i_{(x)jm,k} - \Gamma^i_{(x)jk,m} + \Gamma^r_{(x)jm}\Gamma^i_{(x)rk}  - \Gamma^r_{(x)jk}\Gamma^i_{(x)rm}.
    \end{split}
\ese 
We can see the antisymmetry in the last two entries immediatley from the above. That is ${\Riem^i}_{jkm} = -{\Riem^i}_{jmk}$.

\mybox{
\textbf{Question}: Does a one-dimensional manifold with connection have curvature? Why?
}

\textbf{Solution}: No, as if it is one-dimensional we only have one $\Gamma$, namely ${\Gamma^1}_{11}$, and if we put this into the above definition, all the only component ${\Riem^1}_{111}$ vanishes and so there is no curvature. 

Geometrically this makes sense if we think about embedding the one-dimensional manifold into a higher dimensional space. We can always `pull' the one-dimensional manifold straight and thus show that it has no intrinsic curvature. 

\subsection{The Round Sphere}

\mybox{
There is a question about finding the components of Riem for given $\Gamma$s. I am not going to write it here to save myself time, but I recommend at least watching the video (available \href{https://gravity-and-light.herokuapp.com/tutorials}{here}, for some reason this video is not on YouTube) for a worked calculation. 
}

\subsection{How Not To Define Parallel Transport}

\mybox{
\textbf{Question}: H\"{a}nschen defines two vectors $X\in T_p\cM$ and $Y\in T_q\cM$ as parallel if 
\bse 
    X^i_{(x)} = Y^i_{(x)}
\ese 
with respect to some chart $(U,x)$ whose domain $U$ contains both $p$ and $q$. 

Prove that this notion of parallelity is ill-defined!
}

\textbf{Solution}: Consider the transformation to another chart with the same domain $(U,y)$, 
\bse 
    X^i_{(y)} = \bigg(\frac{\p y^i}{\p x^j}\bigg)_p X^j_{(x)} = \bigg(\frac{\p y^i}{\p x^j}\bigg)_p Y^j_{(x)} = \bigg(\frac{\p y^i}{\p x^j}\bigg)_p \bigg(\frac{\p x^j}{\p y^k}\bigg)_q Y^k_{(y)},
\ese 
but because $p\neq q$ we cannot, in general, use 
\bse 
    \bigg(\frac{\p y^i}{\p x^j}\bigg)_p \bigg(\frac{\p x^j}{\p y^k}\bigg)_q = \del^i_k,
\ese 
and so 
\bse 
    X^i_{(y)} \neq Y^i_{(y)}.
\ese 

\mybox{
There is then a diagram to show how the above definition of parallelity fails for parallel (as defined in the lecture) vectors around a circle when considered in the Cartesian chart and the polar chart. 

To save myself time I have not drawn the diagrams here, but if you can't picture the images in your head, again go see the video!
}

