\section{Schwarzschild Spacetime}

\subsection{Geodesics In A Schwarzschild Spacetime}

\mybox{
The Schwarzschild metric is given and we are told to use the light hand notation 
\bse 
    t(\lambda) := \big(x^0\circ\gamma)(\lambda),
\ese 
and similarly for $r(\lambda)$, $\theta(\lambda)$ and $\varphi(\lambda)$, where $\gamma:\R\to U$ is some curve. 

\textbf{Question}: Write down the Lagrangian $\cL := g_{ab}\dot{\gamma}^a\dot{\gamma}^b$! 
}

\textbf{Solution}: Using the metric given in the question (see the video if you don't know it) we have 
\bse 
    \cL = \bigg(1-\frac{2GM}{r}\bigg) \dot{t}^2 - \bigg(1-\frac{2GM}{r}\bigg)^{-1} \dot{r}^2 - r^2\dot{\theta}^2 - r^2\sin^2\theta \dot{\varphi}^2.
\ese 

\mybox{
\textbf{Question}: Find the Euler-Lagrange equation with respect to $t(\lambda)$!
}

\textbf{Solution}: We see straight away that 
\bse 
    \frac{\p \cL}{\p t} = 0.
\ese 
We also have 
\bse 
    \frac{d}{d\lambda}\bigg(\frac{\p \cL}{\p \dot{t}}\bigg) = 2\bigg(1-\frac{2GM}{r}\bigg) \ddot{t} + \frac{4GM}{r^2}\dot{r}\dot{t},
\ese 
which gives the Euler-Lagrange equation 
\bse 
    \ddot{t} + \frac{2GM}{r^2\Big(1-\frac{2GM}{r}\Big)}\dot{r}\dot{t} = 0
\ese 

\mybox{
\textbf{Question}: Show that the Lie derivative of $g$ with respect to the vector field $K_t := \frac{\p}{\p t}$ vanishes. What does this mean?
}

\textbf{Solution}: We have 
\bse 
    (\cL_{K_t} g)_{ab} = K_t\la g_{ab}\ra + g_{mb}\frac{\p}{\p x^a}(K_t)^m + g_{ab}\frac{\p}{\p x^b}(K_t)^m = 0,
\ese 
as all three terms vanish. This tells us that $K_t$ is a symmetry of the metric. Indeed the Schwarzschild spacetime is stationary (and even static), the definitions for which are given in lecture 16. We will see this symmetry is the conservation of energy.

\mybox{
\textbf{Question}: The exact form of the conserved quantity is given by $(K_t)_a(x^a)'(\lambda)=$const. (without proof). Derive an expression for the quantity $t'(\lambda)$ appearing in the Lagrangian!
}

\textbf{Solution}: Using $(K_t)_a := g_{ab}(K_t)^b$, we have 
\bse 
    g_{ab}(K_t)^b(x^a)'(\lambda) = g_{00}t'(\lambda) = \bigg(1-\frac{2GM}{r}\bigg) t'(\lambda) = \text{const.}
\ese 
Letting $\sqrt{E}$ be the constant, we have 
\bse 
    t'(\lambda) = \frac{r\sqrt{E}}{r-2GM}
\ese 

\mybox{
\textbf{Question}: Moreover, we can find so-called "spherical symmetry", that is, the Lie derivative of $g$ with respect to the already known vector fields 
\bse 
    \begin{split}
        X_1 & = \sin\varphi\frac{\p}{\p\theta} + \cot\theta\cos\varphi\frac{\p}{\p\varphi} \\
        X_2 & = \cos\varphi\frac{\p}{\p\theta} - \cot\theta\sin\varphi\frac{\p}{\p\varphi} \\
        X_3 & = \frac{\p}{\p\varphi}
    \end{split}
\ese 
vanishes. What physical quantity is conserved by this symmetry?
}

\textbf{Solution}: Angular momentum. 

\mybox{
\textbf{Question}: Due to $X_1$ and $X_2$ (without proof), one can fix the motion to a plane of constant $\theta=\frac{\pi}{2}$. How can you derive an expression for the remaining term $\varphi'(\lambda)$?
}
\textbf{Solution}: From two questions above, we have 
\bse 
    \text{const.} = g_{ab}X_3^b(x^a)'(\lambda) = g_{33}\varphi'(\lambda) = -r^2\sin^2\theta\varphi'(\lambda),
\ese
so using $\theta=\frac{\pi}{2}$ and labelling the constant $J$, we have 
\bse 
    \varphi'(\lambda) = \frac{J}{r^2}.
\ese 

\mybox{
\textbf{Question}: Use all the fact that $\cL=1$ on the parameterisation. Insert the previously obtained results and take all terms not containing $E$ to one side!
}

\textbf{Solution}: First note that we have replaced the dots with primes, and also note that $\theta'=0$ as $\theta$ is a constant. We therefore have 
\bse 
    1 = \bigg(1-\frac{2GM}{r}\bigg)\frac{E}{\Big(1-\frac{2GM}{r}\Big)^2} - \frac{1}{1-\frac{2GM}{r}}(r')^2 - r^2\frac{J^2}{r^4},
\ese 
which can be rearranged to 
\bse 
    E = (r')^2 + 1 - \frac{2GM}{r} + \frac{J^2}{r^2} - \frac{2GMJ^2}{r^3}.
\ese 

\mybox{
\textbf{Question}: Can you interpret the terms appearing in this expression?
}
\textbf{Solution}: If we then consider a particle of mass $m=1$, we see the above formula as representing 
\begin{itemize}
    \item $E$ is total energy, 
    \item $(r')^2$ is the kinetic energy, 
    \item $1$ is the mass, 
    \item $-\frac{2GM}{r}$ is the Newtonian gravitational potential, 
    \item $\frac{J^2}{r^2}$ is some angular momentum contribution, and 
    \item $-\frac{2GMJ^2}{r^3}$ is some GR correction term. 
\end{itemize}

\subsection{Gravitational Redshift}

\mybox{
Consider a spacetime equipped with the Schwarzschild metric as well as two observers 1 and 2 at rest in their respective system of reference ($\dot{r}=0$, $\dot{\theta}=0$, $\dot{\varphi}=0$). The observers sit at the same $\theta$ and $\varphi$ while $r_1<r_2$.

\textbf{Question}: Derive an expression for $t'(\lambda)$ using the Lagrangian from the previous exercise!
}

\textbf{Solution}: Using $1=\cL =g_{ab}\dot{\gamma}^a\dot{\gamma}^b$, we get 
\bse 
    t'(\lambda) = \bigg(1-\frac{2GM}{r}\bigg)^{-1/2}.
\ese 

\mybox{
\textbf{Question}: Observer 1 emits photons that observer 2 detects. The gap between the two photon emissions is $\Delta \lambda_1$. Find the gap $\Delta \lambda_2$ seen by observer 2!
}

\textbf{Solution}: We have 
\bse 
    \begin{split}
        \Delta t_1 & = \bigg(1-\frac{2GM}{r_1}\bigg)^{-1/2} \Delta \lambda_1 \\
        \Delta t_2 & = \bigg(1-\frac{2GM}{r_2}\bigg)^{-1/2} \Delta \lambda_2.
    \end{split}
\ese 
Then, we use the fact that $K_t$ was a Killing vector field, and so the path taken by the two emitted photons is the same, apart from a constant time shift.
\begin{center}
    \btik 
        \draw[thick,->] (-0.2,0) -- (5,0);
        \node at (4.8,-0.2) {$r$};
        \draw[thick,->] (0,-0.2) -- (0,3);
        \node at (-0.2,2.8) {$t$};
        \draw[thick] (1,-0.2) -- (1,3);
        \node at (0.8,-0.2) {$r_1$};
        \draw[thick] (4,-0.2) -- (4,3);
        \node at (3.8,-0.2) {$r_2$};
        \draw[thick, blue, decoration={markings, mark=at position 0.5 with {\arrow{>}}}, postaction={decorate}] (1,0.5) .. controls (2,0.6) and (3,0.6) .. (4,1);
        \draw[thick, fill=black] (1,0.5) circle [radius=0.05cm];
        \draw[thick, fill=black] (4,1) circle [radius=0.05cm];
        \draw[thick, blue, decoration={markings, mark=at position 0.5 with {\arrow{>}}}, postaction={decorate}] (1,2) .. controls (2,2.1) and (3,2.1) .. (4,2.5);
        \draw[thick, fill=black] (1,2) circle [radius=0.05cm];
        \draw[thick, fill=black] (4,2.5) circle [radius=0.05cm];
        \node at (0.6,1.25) {\large{$\Delta t_1$}};
        \node at (4.5,1.75) {\large{$\Delta t_2$}};
    \etik 
\end{center}

As the above diagram shows, the Killing condition basically tells us that $\Delta t_1 = \Delta t_2$, and so we get 
\bse 
    \Delta \lambda_2 = \sqrt{\frac{1-\frac{2GM}{r_2}}{1-\frac{2GM}{r_1}}} \Delta \lambda_1
\ese 

\mybox{
\textbf{Question}: Consider the ratio of frequencies $\frac{\omega_1}{\omega_2}$. What happens for observer 2 being approximately at infinity? What happens when sending $r_1$ to the Schwarzschild radius $r_s=2GM$?
}

\textbf{Solution}: We can think of $\Delta \lambda$ being the time period (time between photons emitted) and so the frequency (which is the reciprocal of the time period) is 
\bse 
    \frac{\omega_1}{\omega_2} = \frac{\Delta \lambda_2}{\Delta \lambda_1} = \sqrt{\frac{1-\frac{2GM}{r_2}}{1-\frac{2GM}{r_1}}}.
\ese 
As $r_1<r_2$, the above tells us that $\omega_2<\omega_1$. In terms of wavelength, this is $\mu_1<\mu_2$ (we have used $\mu$ for wavelength as $\lambda$ is already used above), which tells us that the light has been \textit{red-shifted}. This is seen also in the definition of redshift, 
\bse 
    1+z = \frac{\omega_1}{\omega_2},
\ese 
where $z>0$ is redshift and $z<0$ is blueshift. 

For $r_2\to \infty$ we have 
\bse 
    \frac{\omega_1}{\omega_2} \to  \bigg(1-\frac{2GM}{r_1}\bigg)^{-1/2}.
\ese 
For $r_1\to r_s$, we have 
\bse 
    \frac{\omega_1}{\omega_2} \to \infty, 
\ese 
which tells us $\omega_1\to \infty$. In terms of wavelengths, this says $\mu_2\to\infty$, and so the light is infinitely redshifted. This is a so-called \textit{infinite redshift surface}.

This result is actually misleading as it really only holds because of the choice of reference frame. We took the coordinate time to be that of the black hole's rest frame. If we use an in-falling observer's clock to define the coordinate time, this infinite redshift behaviour disappears. I have discussed this in more detail in the notes I have put on my blog site. 

