\section{Relativistic Spacetime, Matter \& Gravitation}

\subsection{Lorentz Force Law}

\mybox{
\textbf{Question}: Recall from the lecture that for a particle coupling to the electromagnetic potential, we have 
\bse 
    m\big(\nabla_{v_{\gamma}}v_{\gamma}\big)^a = q{F^a}_b{v_{\gamma}}^b,
\ese 
where $v_{\gamma}$ is the velocity of a particle of mass $m$ and charge $q$.

Now "$1+3$"-decompose this equation in components with respect to the frame of an observer.
}

\textbf{Solution}: The observer $(\del,e)$ has a frame such that 
\bse 
    g(e_a,e_b) = \eta_{ab}, \qand e_0(\lambda) = v_{\del,\del(\lambda)}.
\ese 
We thus have 
\bse 
    m\big(\nabla_{v_{\gamma}}v_{\gamma}\big)^a = qF^{ac}\eta_{cb}{v_{\gamma}}^b,
\ese 
and so 
\bse 
    \begin{split}
        m\big(\nabla_{v_{\gamma}}v_{\gamma}\big)^0 & = qF^{00}{v_{\gamma}}^0 + \sum_{\beta=1}^3 qF^{0\beta}{v_{\gamma}}^{\beta} \\
        m\big(\nabla_{v_{\gamma}}v_{\gamma}\big)^{\a} & = -qF^{\a0}{v_{\gamma}}^0 - \sum_{\beta=1}^3 qF^{\a\beta}{v_{\gamma}}^{\beta}.
    \end{split}
\ese 


\mybox{
\textbf{Question}: Using the definitions $E_{\a}$ := $F_{\a0}$ for the electric field and $B^{\a} := \frac{1}{2}\varepsilon^{\a\rho\sig}F_{\rho\sig}$ for the magnetic 2
field seen by an observer, bring the right hand side of the above equation to the familiar form of the Lorentz force law for a particle of charge $q$ and spatial velocity
\bse 
    \mathbf{v} := \bigg(\frac{e^{\a}:v_{\del}}{e^0:v_{\del}}\bigg) e_{\a} \qquad (\a=1,2,3\text{ and careful: the denominator was forgotten in lectures})
\ese 
that the observer detects for the particle. 

\textit{Hint: $(a\times b)^{\a} = g^{\a\mu}\varepsilon_{\mu\rho\sig}a^{\rho}b^{\sig}$, $\varepsilon_{123}=1$ and $\epsilon^{123}=1$.}
}

\textbf{Solution}: We set 
\bse 
    \cF^{\a} = m \big(\nabla_{v_{\del}}v_{\del}\big)^{\a} = q{F^{\a}}_b{v_{\del}}^b,
\ese
where $\cF$ is the Lorentz force, up to some factors. We can split the right-hand side into two terms, one for $b=0$ and the other for $b=\beta$. For the former we note that 
\bse 
    \eta^{\a\beta}E_{\beta} = \eta^{\a\beta}F_{\beta0} = {F^{\a}}_0,
\ese 
so, using $\epsilon^0:(v_{\del}) = {v_{\del}}^0$, the first term is simply 
\bse 
    q\eta^{\a\beta}E_{\beta}\big(\epsilon^0:v_{\del}\big).
\ese 
The second term need a little more work, but we start using the hint: using $g^{ab}=\eta^{ab}$ in the observers frame, we have
\bse 
    (v_{\del}\times B)^{\a} = \eta^{\a\mu}\varepsilon_{\mu\rho\sig} {v_{\del}}^{\rho}B^{\sig} := \frac{1}{2}\eta^{\a\mu}\varepsilon_{\mu\rho\sig}\varepsilon^{\sig\nu\tau} {v_{\del}}^{\rho}F_{\nu\tau}. 
\ese 
We see that the $\eta^{\a\mu}$ tells us that $\mu=\a$ in the Levi-Civita symbol. By definition, the only non-vanishing terms on the right, then, have $\rho\neq\sig\neq\a$. Let's consider the case for $\a=1$ (the other two follow analogously)
\bse 
    \begin{split}
        \frac{1}{2}\eta^{1\mu}\varepsilon_{\mu\rho\sig}\varepsilon^{\sig\nu\tau}F_{\nu\tau}{v_{\del}}^{\rho} & = \frac{1}{2}\varepsilon_{123}\big(\varepsilon^{312}F_{12} + \varepsilon^{321}F_{21}\big){v_{\del}}^2 + \frac{1}{2}\varepsilon_{132}\big(\varepsilon^{213}F_{13} + \varepsilon^{231}F_{31}\big){v_{\del}}^3 \\
        & = \frac{1}{2} \varepsilon_{123} \big( \varepsilon^{123}F_{12} + (-\varepsilon^{123})(-F_{12})\big) {v_{\del}}^2 + \frac{1}{2} (-\varepsilon_{123}) \big( (-\varepsilon^{123})F_{13} + \varepsilon^{123}(-F_{13})\big){v_{\del}}^3 \\
        & = F_{12}{v_{\del}}^2 + F_{13}{v_{\del}}^3,
    \end{split}
\ese 
where we have indicated where the antisymmetries of $\varepsilon$ and $F$ have been used, and we have also used $\varepsilon_{123}=1=\varepsilon^{123}$. This generalises to 
\bse 
    \frac{1}{2}\eta^{\a\mu}\varepsilon_{\mu\rho\sig}\varepsilon^{\sig\nu\tau}F_{\nu\tau}{v_{\del}}^{\rho} = \eta^{\a\sig}F_{\sig\beta}{v_{\del}}^{\beta} = {F^{\a}}_{\beta}{v_{\del}}^{\beta},
\ese
where we note the fact that $F_{\a\a}=0$ due to antisymmetry. 

So the second term in our Lorentz force equation is simply 
\bse 
    q{F^{\a}}_{\beta}{v_{\del}}^{\beta} = q\big( v_{\del} \times B \big)^{\a},
\ese 
giving us 
\bse 
    \cF^{\a} = q\eta^{\a\beta} E_{\beta} \big( \epsilon^0:v_{\del}\big) + q\big(v_{\del}\times B\big)^{\a}.
\ese 
Finally, we note that  
\bse 
    \frac{1}{(\epsilon^0:v_{\del})} \big(v_{\del}\times B\big)^{\a} = \big(\mathbf{v}\times B\big)^{\a},
\ese 
as everywhere in the calculation above we'll get terms like 
\bse 
    \frac{v_{\del}^\rho}{(\epsilon^0:v_{\del})} = \frac{(\epsilon^\rho:v_{\del})}{(\epsilon^0:v_{\del})} =: \mathbf{v}^{\rho}.
\ese 
This gives us the form we want on the right-hand side, i.e.
\bse 
    \frac{1}{(\epsilon^0:v_{\del})}\cF^{\a} = q\eta^{\a\beta}E_{\beta} + q\big(\mathbf{v}\times B\big)^{\a}. 
\ese 

\subsection{Which Curvature Can Feature In Einstein’s Equations?}

\mybox{
This questions asks to show that the differential Bianchi identity holds for the Riemann tensor. This was given as an exercise in the lecture. 

If the reader couldn't work it out there, \href{https://math.stackexchange.com/questions/1494262/direct-proof-of-the-second-bianchi-identity}{this link} should prove helpful. Note the notation is slightly different in the link, but of course the answer is the same.

I didn't include this link in the notes to hopefully avoid temptation of just looking up the answer.
}

\mybox{
\textbf{Question}: The above component-free version can equivalently be written as 
\bse 
    {R^w}_{zab;c} + {R^w}_{zbc;a} + {R^w}_{zca;b} = 0.
\ese
Using this result, show that by appropriate contractions one obtains
\bse 
    (\nabla_aG)^{ab}=0.
\ese 
}

\textbf{Solution}: First use the antisymmetry to give 
\bse 
    {R^w}_{zab;c} + {R^w}_{zbc;a} + {R^w}_{zca;b} = {R^w}_{zab;c} + {R^w}_{zbc;a} - {R^w}_{zac;b}.
\ese 
Now, we can use the fact that $\nabla$ is metric-compatible and the result 
\bse 
    g^{av}g_{vw} = \del^a_w
\ese
to give
\bse 
    \begin{split}
        g^{av}g_{vw}\big({R^w}_{zab;c} + {R^w}_{zbc;a} + {R^w}_{zca;b}\big) & = {R^a}_{zab;c} + {R^a}_{zbc;a} - {R^a}_{zac;b} \\
        & = R_{zb;c} - R_{zc;b} + {R^a}_{zbc;a}.
    \end{split}
\ese 
Then contract with $g^{bz}$ along with the results $R^{\,\,\,a}_{z\,\,\,bc} = -{R^a}_{zbc}$\footnote{We essentially showed this result in tutorial 9 (we showed $R_{abcd}=-R_{bacd}$).} and $R:= g^{ab}R_{ab}$ to give 
\bse 
    \begin{split}
        g^{bz}\big(R_{zb;c} - R_{zc;b} + {R^a}_{zbc;a}\big) & = R_{;c} - {R^b}_{c;b} - {R^{ba}}_{bc;a} \\
        & = R_{;c} - {R^b}_{c;b} - {R^a}_{c;a} \\
        & = R_{;c} - 2{R^b}_{c;b}.
    \end{split}
\ese 
Finally using the fact that the Ricci tensor is symmetric,\footnote{This result is obtained from $R_{abcd}=R_{cdab}$, then setting $a=c$ and raising the first index back up.} and contracting with $g^{ac}$, we have (after relabelling)
\bse 
    \big(\nabla_aG)^{ab} = {R^{ab}}_{;a} - \bigg(\frac{1}{2}g^{ab}R\bigg)_{;a} = 0.
\ese 

