\section{Tangent Spaces}

\subsection{Virtuoso Use Of The Symbol $\big(\frac{\p}{\p x^i}\big)_p$}

\mybox{
\textbf{Question}: Show that, for overlapping charts $(U,x)$ and $(V,y)$, one has 
\bse 
    \bigg(\frac{\p x^a}{\p y^m}\bigg)_p \bigg(\frac{\p y^m}{\p x^b}\bigg)_p = \del^a_b
\ese
for any $p\in U\cap V$. 
}
\textbf{Solution}: We have 
\bse 
    \begin{split}
        \del^a_b & = \bigg(\frac{\p}{\p x^b}\bigg)_p(x^a) \\
        & := \p_b\big(x^a\circ x^{-1}\big)\big(x(p)\big) \\
        & = \p_b\big(x^a\circ (y^{-1}\circ y) \circ x^{-1}\big)\big(x(p)\big) \\
        & = \p_b\big((x^a\circ y^{-1}) \circ (y \circ x^{-1})\big)\big(x(p)\big) \\
        & = \p_b\big(y^m\circ x^{-1}\big)\big(x(p)\big) \cdot \p_m\big(x^a\circ y^{-1}\big) \big(y(p)\big) \\
        & =:  \bigg(\frac{\p y^m}{\p x^b}\bigg)_p \bigg(\frac{\p x^a}{\p y^m}\bigg)_p,
    \end{split}
\ese 
where we have inserted the identity, used the associativity of the composition of maps, and the multidimensional chain rule along with $\big(y\circ x^{-1}\big)\big(x(p)\big) = y(p)$. 

\mybox{
\textbf{Question}: After inserting $y^{-1}\circ y$, where $y$ is another chart map on the same chart domain $U$, at the appropriate position in the definition of the left hand side of 
\bse 
    \bigg(\frac{\p f}{\p x^i}\bigg)_p = \bigg(\frac{\p y^m}{\p x^i}\bigg)_p\bigg(\frac{\p f}{\p y^m}\bigg)_p,
\ese 
use the multidimensional chain rule to show that it equals the right-hand side. 
}
\textbf{Solution}: This follows in a similar manner to the previous question and so I won't type it here. 

\mybox{
\textbf{Question}: Do the $\dim\cM$ many quantities defined by the left-hand side of the above expression constitute the components of a tensor? If so, what is the valence and rank of the tensor?
}
\textbf{Solution}: The above expression is clearly of the form of 
\bse 
    T_{(x)i}(p) = \bigg(\frac{\p y^m}{\p x^i}\bigg)_p T_{(y)m}(p),
\ese 
which is the transformation law for the components of a $(0,1)$-tensor. So the answer is "yes" and the valence is $(0,1)$ and the rank is $1$. 

\subsection{Transformation Of Vector Components}

\mybox{
Let the topological space $(\R^2,\cO_{st.})$ be equipped with the atlas $\cA = \{(\R^2,x),(\R^2,y)\}$, where 
\bse 
    x : (a,b) \mapsto (a,b) \qand y:(a,b) \mapsto (a,b+a^3).
\ese 
\textbf{Question}: Calculate the objects $\big(\frac{\p x^i}{\p y^j}\big)_p$!
}

\textbf{Solution}: We have $\big(x \circ y^{-1}\big)\big((u,v)\big) = (u, v-u^3)$, and so direct calculation gives 
\bse 
    \bigg(\frac{\p x^1}{\p y^1}\bigg)_p := \p_1 \big(x^1 \circ y^{-1}\big)\big(y(p)\big) = 1,
\ese 
where we have used $\p_1\big(x^\circ y^{-1}\big)\big((u,v)\big) = 1$, irrespective of the value $u$ takes.  

Next we have,
\bse 
    \bigg(\frac{\p x^2}{\p y^1}\bigg)_p := \p_1\big( x^2 \circ y^{-1}\big)\big(y(p)\big) = \p_1\big( x^2 \circ y^{-1}\big)\big((a,b+a^3)\big) = -3a^2,
\ese 
where we have used $\p_2\big(x^1\circ y^{-1}\big)\big((u,v)\big) = -2u^2$ along with $p=(a,b)$. 

Similar calculations give 
\bse 
    \bigg(\frac{\p x^1}{\p y^2}\bigg)_p = 0, \qand \bigg(\frac{\p x^2}{\p y^2}\bigg)_p = 1
\ese

\mybox{
(Reworded slightly to save typing) Recall that the components of the velocity to a curve $\gamma$ in a chart $(U,x)$ at point $p=\gamma(\lambda_0)$ are given by 
\bse 
    \dot{\gamma}_x^i(\lambda_0) := \big(x\circ \gamma)^{i\prime} (\lambda_0).
\ese 
Now consider the curve 
\bse 
    \gamma :\R\to \R^2; \quad  \lambda\mapsto (\lambda,-\lambda). 
\ese
\textbf{Question}: Calculate the components $\dot{\gamma}_x^i (\lambda_0)$ and $\dot{\gamma}_y^i (\lambda_0)$!
}

\textbf{Solution}: We have $(x\circ\gamma)(\lambda) = (\lambda, -\lambda)$ and $(y\circ\gamma)(\lambda) = (\lambda, -\lambda+\lambda^3)$, so we have 
\bse 
    \begin{split}
        \dot{\gamma}_x^1(\lambda_0) & := \big(x\circ \gamma)^{1\prime} (\lambda_0) = 1, \\
        \dot{\gamma}_x^2(\lambda_0) & := \big(x\circ \gamma)^{2\prime} (\lambda_0) = -1, \\
        \dot{\gamma}_y^1(\lambda_0) & := \big(y\circ \gamma)^{1\prime} (\lambda_0) = 1, \\
        \dot{\gamma}_y^2(\lambda_0) & := \big(x\circ \gamma)^{2\prime} (\lambda_0) = -1+3\lambda_0^2. 
    \end{split}
\ese 

\mybox{
\textbf{Question}: With the above results in mind, how could you have obtained the components of $\dot{\gamma}_x^i(\lambda_0)$ from the $\dot{\gamma}_y^i(\lambda_0)$?
}

\textbf{Solution}: The answer is clearly to just use the transformation property 
\bse 
    \dot{\gamma}_x^i(\lambda_0) = \bigg(\frac{\p x^i}{\p y^m}\bigg)_p \dot{\gamma}_y^m(\lambda_0).
\ese 
That is, we have
\bse 
    \begin{split}
        \dot{\gamma}_x^1(\lambda_0) & = \bigg(\frac{\p x^1}{\p y^1}\bigg)_p \dot{\gamma}_y^1(\lambda_0) + \bigg(\frac{\p x^1}{\p y^2}\bigg)_p \dot{\gamma}_y^2(\lambda_0) \\
        & = 1\cdot 1 + 0\cdot(-1+3\lambda_0)^2 \\
        & = 1,
    \end{split}
\ese 
and 
\bse 
    \begin{split}
        \dot{\gamma}_x^2(\lambda_0) & = \bigg(\frac{\p x^2}{\p y^1}\bigg)_p \dot{\gamma}_y^1(\lambda_0) + \bigg(\frac{\p x^2}{\p y^2}\bigg)_p \dot{\gamma}_y^2(\lambda_0) \\
        & = -3\lambda_0^2 \cdot 1 + 1\cdot (-1+3\lambda_0^2) \\
        & = -1.
    \end{split}
\ese 

\subsection{The Gradient}

\mybox{
Given a function $f$ on a manifold $\cM$, the level sets of $f$ for a constant $c\in\R$ are defined as 
\bse 
    N_c(f) := \{p\in\cM \, | \, f(p) = c\}. 
\ese 
\textbf{Question}: Formulate the condition for a curve $\gamma:\R\to\cM$ to take values solely in one of the level sets of a function $f$!
}
\textbf{Solution}: Clearly we just want $(f\circ \gamma)(\lambda)=c$ for all $\lambda\in\R$. We can formulate this as $f\circ \gamma$ being equivalent to the constant function, which simply obeys the above property. 

\mybox{
\textbf{Question}: Now show that the gradient of the function annihilates the velocity vector $v_{\gamma,p}$ for any such $\gamma$ through $p$ in $N_c(f)$. In other words, show that 
\bse 
    (df)_p (v_{\gamma,p})=0. 
\ese
}
\textbf{Solution}: We have 
\bse 
    (df)_p (v_{\gamma,p}) := v_{\gamma,p}(f) := \big(f\circ \gamma)^{\prime}(\lambda_0) =0,
\ese 
as the derivative of a constant vanishes. 

\subsection{Is There A Well-Defined Sum Of Curves?}

\mybox{
Let the topological manifold $(\R^2,\cO_{st.})$ be equipped with the atlas $\cA=\{(\R^2,x),(\R^2,y)\}$ where 
\bse 
    x:(a,b) \mapsto (a,b) \qand y:(a,b)\mapsto (a, b\cdot e^a).
\ese 
\textbf{Question}: Is $\cA$ a $C^{\infty}$-atlas?
}
\textbf{Solution}: We need to construct the chart transition maps 
\bse 
    \begin{split}
        (x\circ y^{-1})(u,v) & = x\big((u, v \cdot e^{-u})\big) = (u,v\cdot e^{-u}) \\
        (y\circ x^{-1})(u,v) & = y\big((u,v)\big) = (u, v\cdot e^u). 
    \end{split}
\ese 
These are both infinitely times continuously differentiable and so the answer is "yes". 

\mybox{
\textbf{Question}: On our $\R^2$ above, consider two curves $\gamma,\del:\R\to\R^2$ given by 
\bse 
    \gamma: \lambda \mapsto (\lambda, 1),  \qand  \del :\lambda \mapsto  (1,\lambda).
\ese 
Without referring to any chart, can you give the sum $\gamma + \del$ of these curves? 
}
\textbf{Solution}: The answer is "no" because our manifold is just the set $\R^2$ (with a topology) and so carries no vector space structure so we cannot talk about the addition on $\R^2$. 

\mybox{
\textbf{Question}: Calculate the representatives of both curves with respect to both charts. Illustrate the results. Where do the curves in the charts intersect?
}

\textbf{Solution}: First consider the $x$ chart. We have 
\bse 
    (x\circ \gamma) (\lambda) = (\lambda, 1), \qand (x\circ \del) (\lambda) = (1,\lambda).
\ese 

In the $y$ chart we have 
\bse 
    (y\circ \gamma) (\lambda) = (\lambda, e^{\lambda}), \qand (y\circ \del) (\lambda) = (1,\lambda\cdot e)
\ese 
The illustrations are as follows 

\begin{center}
    \btik 
        \draw[thick,->] (-3,0) -- (3,0);
        \draw[thick,->] (0,-2) -- (0,3);
        \node at (2.8,-0.5) {\large{$x^1$}};
        \node at (-0.5,1.8) {\large{$x^2$}};
        \draw[ultra thick, blue] (-3,1) -- (3,1);
        \draw[ultra thick, red] (1,-2) -- (1,3);
        \node at (-2,1.2) {\large{\textcolor{blue}{$\gamma$}}};
        \node at (1.2,2) {\large{\textcolor{red}{$\del$}}};
        % 
        \draw[thick,->] (5,0) -- (11,0);
        \draw[thick,->] (8,-2) -- (8,3);
        \node at (10.8,-0.5) {\large{$y^1$}};
        \node at (7.5,1.8) {\large{$y^2$}};
        \draw[ultra thick, blue] (5,0.2) .. controls (8.5,0.5) .. (10,2.8);
        \draw[ultra thick, red] (9,-2) -- (9,3);
        \node at (10,2.4) {\large{\textcolor{blue}{$\gamma$}}};
        \node at (8.8,2) {\large{\textcolor{red}{$\del$}}};
    \etik 
\end{center}

The intersection points are $(1,1)$ in the $x$ chart and $(1,e)$ in the $y$ chart. These both correspond to $\lambda=1$, which it must from the definition of the curves. 

\mybox{
(Reworded to save typing)
\textbf{Question}: Use the formula from the lectures 
\bse 
    \sig_x: \R\to U; \quad \lambda \mapsto x^{-1}\big( (x\circ\gamma)(\lambda+\lambda_0) + (x\circ\del)(\lambda+\lambda_1) - (x\circ\gamma)(\lambda_0)\big),
\ese 
where $\gamma(\lambda_0)=\del(\lambda_1)$, to find the sum $\gamma+\del$. Also do the calculation for $\sig_y(\lambda)$.
}

\textbf{Solution}: Using the previous question, direct calculation gives: for the $x$ chart 
\bse 
    \begin{split}
        \sig_x(\lambda) & = x^{-1}\big( (\lambda+1,1) + (1,\lambda+1) - (1,1)\big) \\
        & = x^{-1}(\lambda+1,\lambda+1) \\
        & = (\lambda+1,\lambda+1).
    \end{split}
\ese 
The calculation i the $y$ chart gives 
\bse 
    \sig_y(\lambda) = (\lambda+1, 1+ \lambda \cdot e^{-\lambda}). 
\ese 

\mybox{
\textbf{Question}: Show that -- desipite the above results -- the velocity of $\sig_x$ and the velocity $\sig_y$ are equal at the intersection point.  
}

\textbf{Solution}: The intersection point is $(1,1)$ in $x$ and so we require $\lambda=0$. Using this, we have 
\bse 
    \dot{\sig}_{(x)x}^1 = 1 \qand \dot{\sig}_{(x)x}^2 = 1,
\ese
as both are just the derivative w.r.t. $\lambda$ of $\lambda+1$. 

In the $y$ chart we have (note we use the $x$ chart in to find these components in order to compare it to the above ones)
\bse 
    \dot{\sig}_{(x)y}^1 = 1 \qand \dot{\sig}_{(x)y}^2 = 1 \cdot e^0 - 0\cdot e^0 = 1.
\ese 

So we have 
\bse 
    v_{\sig_x,(1,1)} = \dot{\sig}_{(x)x}^i \bigg(\frac{\p}{\p x^i}\bigg)_p = \dot{\sig}_{(x)y}^i\bigg(\frac{\p}{\p x^i}\bigg)_p = v_{\sig_y,(1,1)}. 
\ese 

