\section{Topological Manifolds}

\subsection{An Atlas From A Real World --- the M\"{o}bius river}
\mybox{
\textbf{Question}: Consider a M\"{o}bius strip\footnote{Please Google it if you don't know what it is.} with a river drawn along it. How many charts do you need to cover the M\"{o}bius strip? 

Draw an image of the river and M\"{o}bius strip under the chat map(s)!
}

\textbf{Solution}: The M\"{o}bius strip can be represented by the following diagram 
\begin{center}
    \btik 
        \draw[thick] (0,0) -- (5,0);
        \draw[thick] (0,2) -- (5,2);
        \draw[thick, decoration={markings, mark=at position 0.5 with {\arrow{>}}}, postaction={decorate}] (0,0) -- (0,2);
        \draw[thick, decoration={markings, mark=at position 0.5 with {\arrow{>}}}, postaction={decorate}] (5,2) -- (5,0);
        \draw[blue, ultra thick] (0,1.5) .. controls (2,0) and (4,1.5) ..(5,0.5);
    \etik 
\end{center}
where the arrows indicate how we identify the two edges together (i.e. the bottom right corner goes to the top left corner). The blue line is the river. It is clear that we will need at least two charts in order to map the whole strip. We could choose them as follows 

\begin{center}
    \btik 
        \draw[dashed, fill=lightgray] (0,0) -- (0,2) -- (1.5,2) -- (1.5,0) -- (0,0);
        \draw[dashed, fill=lightgray] (5,0) -- (5,2) -- (3.5,2) -- (3.5,0) -- (5,0);
        \draw[thick] (0,0) -- (5,0);
        \draw[thick] (0,2) -- (5,2);
        \draw[thick, decoration={markings, mark=at position 0.5 with {\arrow{>}}}, postaction={decorate}] (0,0) -- (0,2);
        \draw[thick, decoration={markings, mark=at position 0.5 with {\arrow{>}}}, postaction={decorate}] (5,2) -- (5,0);
        \draw[blue, ultra thick] (0,1.5) .. controls (2,0) and (4,1.5) ..(5,0.5);
        \node at (0.75,0.5) {\Large{$U_1$}};
        \node at (4.25,1.5) {\Large{$U_1$}};
        %
        \draw[dashed, fill=lightgray] (7.5,0) -- (7.5,2) -- (11.5,2) -- (11.5,0) -- (7.5,0);
        \draw[thick] (7,0) -- (12,0);
        \draw[thick] (7,2) -- (12,2);
        \draw[thick, decoration={markings, mark=at position 0.5 with {\arrow{>}}}, postaction={decorate}] (7,0) -- (7,2);
        \draw[thick, decoration={markings, mark=at position 0.5 with {\arrow{>}}}, postaction={decorate}] (12,2) -- (12,0);
        \draw[blue, ultra thick] (7,1.5) .. controls (9,0) and (11,1.5) ..(12,0.5);
        \node at (9.5,1.5) {\Large{$U_2$}};
        %
    \etik 
\end{center}
We then define maps from these shaded regions into $\R^2$. For example we can just map them so that they are the shaded regions on the page, taking care to region the two parts of $U_1$ properly. To save my writing more Tikz code, I'll leave this to your imagination. 

\subsection{A Real World From An Atlas}

\mybox{
This is worth watching on \href{https://www.youtube.com/watch?v=ghfEQ3u_B6g&list=PLFeEvEPtX_0RQ1ys-7VIsKlBWz7RX-FaL&index=2}{the video}, but would be way too hard for me to do on here. So please go watch the video for this one.
}

\subsection{Before The Invention Of The Wheel}

\mybox{
Consider the set $F^1:= \{(m,n)\in\R^2 \, | \, m^4 + n^4 = 1\}$ of pairs of real numbers $(m,n)$. Let it be equipped with the topology subset topology $\cO_{s}|_{F^1}$ inherited from the standard topology on $\R^2$.  

\textbf{Question}: We look at a map $x:F^1\to \R$ that maps a pair in $F^1$ to the first entry in the pair. Write this in formal mathematical terms! Is the map injective?
}

\textbf{Solution}: The map is $x: F^1 \to [-1,1]$ given by $x:(m,n) 
\mapsto m$. Clearly this map is not injective as $x\big((m,n)\big) = x\big((m,n')\big)$ where $n' = -n$. 

\mybox{
\textbf{Question}: This map may be made injective by restricting its domain to either of two maximal open subsets of $F^1$. Which ones? Call them $x_{\uparrow}$ and $x_{\downarrow}$.
}

\textbf{Solution}: The problem was that $n'=-n$ gave non-injectivity, so clearly we define
\bse 
    \begin{split}
        U_{\uparrow}  & := \{(m,n)\in\R^2 \, | \, m^4 + n^4 = 1, n>0\} \\
        U_{\downarrow}  & := \{(m,n)\in\R^2 \, | \, m^4 + n^4 = 1, n<0\} 
    \end{split}
\ese 
and so have the maps 
\bse 
    x_{\uparrow} : U_{\uparrow} \to (-1,1), \qquad \text{and} \qquad x_{\downarrow} : U_{\downarrow} \to (-1,1).
\ese 
It is important that we take the inequalities for $n$ (i.e. $n\neq 0$) so that our sets are open, as required by the question. 

\mybox{
\textbf{Question}: Now, construct an injective map $y_{\uparrow}:F^1 \to \R$ that maps every pair in a maximal open subset of $F^1$ to the second entry of the pair. 
}

\textbf{Solution}: Same as above we simply define 
\bse 
    V_{\uparrow}  := \{(m,n)\in\R^2 \, | \, m^4 + n^4 = 1, m>0\},
\ese 
and the map 
\bse 
    y_{\uparrow} : V_{\uparrow} \to (-1,1),
\ese 
given by $y_{\uparrow}:(m,n)\mapsto n$. 

\mybox{
\textbf{Question}: Is $y_{\uparrow}$ invertible? If so, construct $y_{\uparrow}^{-1}$!
}

\textbf{Solution}: Yes it is invertible as it is bijective. We construct the inverse as 
\bse 
    y_{\uparrow}^{-1} : (-1,1) \to V_{\uparrow},
\ese 
where the action is given by
\bse 
    y_{\uparrow}^{-1} : n \mapsto \big( \sqrt[4]{1-n^4}, n\big).
\ese 

\mybox{
\textbf{Question}: Do the domains of the maps $x_{\uparrow}$ and $y_{\uparrow}$ overlap? If so, construct the \textit{transition map} $x_{\uparrow}\circ y_{\uparrow}^{-1}$ and specify its domain and target. 
}

\textbf{Solution}: Yes their domains overlap as 
\bse 
    U_{\uparrow}\cap V_{\uparrow} = \{(m,n)\in\R^2 \, | \, m^4 + n^4 = 1, n,m>0\}
\ese 
The transition map is 
\bse 
    \big(x_{\uparrow}\circ y_{\uparrow}^{-1}\big) : (0,1) \to (0,1)
\ese 
with 
\bse 
    \big(x_{\uparrow}\circ y_{\uparrow}^{-1}\big)(a) = x_{\uparrow}\big(\sqrt[4]{1-a^4}, a\big) = \sqrt[4]{1-a^4} \in (0,1).
\ese 

\mybox{
\textbf{Question}: How many maps (constructed this way) do you need for their domains to cover the whole of $F^1$? Does the collection of these domains and maps form an atlas on $F^1$?
}

\textbf{Solution}: The answer is obviously 4 maps, $x_{\uparrow},x_{\downarrow},y_{\uparrow}$ and $y_{\downarrow}$, and yes they form an atlas. From this we see that $F^1$ is a topological manifold of dimension $d=1$.

